---
title: 'From R to SVG: Our workflow to generate force-layout networks'
author: "Jan Sodoge"
date: "8/30/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```


This file documents the process on the R-level described in the compscjournalism.org tutorial on getting from network data in R to a force-layout network as an SVG in Javascript and D3. 

#Getting the data ready in R


```{r}
library(statnet)
library(tidyverse)
library(knitr)
load("sample_data.RData") #<- find this sample data attached in the repository
```
The sample data we start with consist of a vector containing the nodes (i.e. basketball player names) and a dataframe, containing the ties between them (see original formatting below).
```{r}
kable(head(edges,3))
```





### Create the nodelist
```{r}

#we create the nodelist within the format required
nodelist_d3 <- as.data.frame(nodes) #<- since the nodes are stored in vector format
nodelist_d3$group <- 1 # if you e.g.  wish to color nodes by group, pass a groupID here, in this example all groups identify with one group

nodelist_d3$ID <- seq(0, nrow(nodelist_d3)-1) #<- get each player an individual ID creating the edgelist further on
colnames(nodelist_d3) <- c("name", "group", "ID")
kable(head(nodelist_d3,3))

```



### Create the edgelist


```{r}
edgelist_d3 <- edges #edges as the sample data dataframe
edgelist_d3$weight <- 1 # add a weight to the edges
colnames(edgelist_d3) <- c("source", "target", "value")
```

Next, we need to replace the player names with their respective IDs created previously within the nodelist
```{r}
edgelist_id_version <- merge(edgelist_d3, nodelist_d3, by.x = "source", by.y = "name") %>% 
  select(ID, target, value)
#merge for target
edgelist_id_version <- merge(edgelist_id_version, nodelist_d3, by.x = "target", by.y = "name") %>% 
  select(ID.x, ID.y, value)
colnames(edgelist_id_version) <- c("source", "target", "value")
edgelist_d3 <- edgelist_id_version

```



Thus, our data we will pass on to Javascript looks like this. For the edgelist, the numbers indicate the IDs of the nodes (i.e. players). 
```{r}
kable(head(nodelist_d3,3))
kable(head(edgelist_d3,3))

```



### Getting the data into JSON format
We use the d3_json function from d3r package to create a JSON object based on the two previously created dataframes. Later, the .html file containing the Javascript code will load this file to visualize the network. 
```{r}
library(RJSONIO)
library(d3r)
write(paste0("{", d3_json(list(nodes = nodelist_d3, links = edgelist_d3)), "}"),
        "graph_data.json")
```



### That's it on the R-level
Having created the .json file, R has finished its duties. From now, we can focus on the JS/D3/HTML side. To follow along, switch back to the main tutorial overview. 